MIMO Channel Shaping and Rate Maximization\\Using Beyond Diagonal RIS

### 2025-02-25 review/response_241203.tex

\end{editor}

\begin{reviewer}
	\summary{
		This manuscript investigates the impact of \gls{bd}-\gls{ris} to the distribution of singular values of \gls{mimo} channel of a point-to-point communication system. Especially, the authors derive bounds of singular values of \gls{bd}-\gls{ris} channel and propose algorithms to design \gls{bd}-\gls{ris} configuration.
	}

	\comment{
		\label{cm:a}
		The contribution of Sec. \ref{sc:shaping_numerical} is unclear for the following two reasons.
		\begin{enumerate}
			\item The manuscript claims to propose a universal optimization framework for \gls{bd}-\gls{ris} as in \eqref{op:shaping}. However, the framework in \eqref{op:shaping} is indeed not sufficiently general. It does not consider \gls{bd}-\gls{ris} dependent constraints (e.g., \gls{qos}, sensing quality) besides the orthogonal conditions \eqref{cn:shaping_unitary}. The reviewer notice that a more general framework, which considers \gls{bd}-\gls{ris} constraints in addition to \eqref{cn:shaping_unitary}, has already been investigated and solved in the authors' previous work \cite{Zhou2023}.
			\item Besides, both the non-geodesic \gls{rcg} and geodesic \gls{rcg} algorithms, e.g., \eqref{eq:gradient_eucl} -- \eqref{eq:geodesic_tran}, are well established methods in the literature, e.g., \cite{Abrudan2008}. Therefore, the novelty of Sec. \ref{sc:shaping_numerical} unclear.
		\end{enumerate}
	}

	\reply{
		We appreciate the reviewer's insightful comments.
		The geodesic \gls{rcg} algorithm for unitary-constrained problems has indeed been proposed in the optimization literature.
		Our contribution is an extension to block-unitary cases with parallel or unified updates that accelerate the design of group-connected \gls{bd}-\gls{ris}.
		The claim of ``universal optimization framework'' and the non-geodesic \gls{rcg} method have been removed and we have completely reworked the optimization part to reflect the particular contributions (channel shaping).
		The content is too long to fit here and we kindly ask the reviewer to refer to Section \ref{sc:shaping_numerical} of the revised manuscript for details.
	}

@@

	\comment{
		In Sec. \ref{sc:shaping_numerical}, the manuscript proposes to maximize weighted sum of singular values of the effective channel $\mathbf{H}$. The optimization objective is rather vague. Is the weighted sum of singular values related to any meaningful performance metric (e.g., capacity, power gain, and so on)?
	}

	\reply{
		This is an important question.
		The weighted sum of singular values plays a crucial role in channel shaping since it helps to reveal the achievable singular value region.
		We believe itself could serve as a metric for measuring the \gls{mimo} channel shaping capability of different \gls{ris}.
		We have added a detailed explanation in the revised manuscript and the answer which we copy in the two boxes below.

@@

	\comment{
		\label{cm:b1}
		In Sec. \ref{sc:shaping_analytical}, the manuscript provides a number of bounds on singular values of the effective channel $\mathbf{H}$. However, the usefulness of these bounds is unclear. What can these bounds be used for? Note the results in Corollaries \ref{co:nd_sv_prod_subset} -- \ref{co:nd_sv_prod_tail} are quite standard results from matrix textbook. The (upper) bound in Corollary \ref{co:nd_power}, which is achieved by aligning its left and right singular-value vectors to that of the forward and backward channels, respectively, is reminiscent of the results in seminal papers on relay beamforming design, e.g., \cite{Rong2009,Rong2009a}.
	}

	\reply{
		We would like to argue that, while Corollaries \ref{co:nd_sv_prod_subset} -- \ref{co:nd_sv_prod_tail} are standard mathematical results, it is Proposition \ref{pp:nd} that allows one to recast the channel shaping question as a well-studied linear algebra problem and enable those.
		In the revised manuscript, we have emphasized above and have also added

### 2025-02-25 manuscript/appendix.tex
:construction: Merge footnotes

\input{preamble}

\title{MIMO Channel Shaping and Rate Maximization\\Using Beyond Diagonal RIS}
\author{
	\IEEEauthorblockN{
		Yang Zhao, \IEEEmembership{Member, IEEE,}
		Hongyu Li, \IEEEmembership{Graduate Student Member, IEEE,}\\
		Massimo Franceschetti, \IEEEmembership{Fellow, IEEE,}
		and Bruno Clerckx, \IEEEmembership{Fellow, IEEE}
	}
	\thanks{
		Yang Zhao is with the School of Electronic and Electrical Engineering, University of Leeds, LS2 9JT, U.K. (e-mail: \href{mailto:yang.zhao@ieee.org}{yang.zhao@ieee.org}.)

		Hongyu Li and Bruno Clerckx are with the Department of Electrical and Electronic Engineering, Imperial College London, SW7 2AZ, U.K.

		Massimo Franceschetti is with the Department of Electrical and Computer Engineering, University of California, San Diego, La Jolla CA 92093, USA.

		This work has been partially supported by UKRI grant EP/Y004086/1, EP/X040569/1, EP/Y037197/1, EP/X04047X/1, EP/Y037243/1.
	}
}

@@

\maketitle

\begin{abstract}
	This paper investigates the limits to which a passive \gls{ris} can reshape a point-to-point \gls{mimo} in terms of singular values
	% and power gain
	and thus improve the rate performance.
	% redistribute the singular values of a point-to-point \gls{mimo} channel for providing power and rate gains.
	We depart from the conventional \gls{ris} with \gls{d} phase shift matrix and adopt a \gls{bd} model interconnecting elements for signal amplitude and phase manipulation.
	This translates to superior branch matching and mode alignment capabilities in \gls{mimo}.
	% featuring group-wise connections for enhanced branch matching and mode alignment capabilities.
	Specifically, analytical singular values bounds are derived under popular \gls{ris} deployment scenarios
	% the scenario of rank-deficient backward/forward channel or negligible direct channel, which
	to showcase the shaping advantages of \gls{bd}-\gls{ris} in terms of dynamic range and \gls{dof}.
	A numerical optimization method is then proposed for a broader range of shaping problems and invoked to characterize the achievable singular value region.
	As a side product, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem by a local-optimal \gls{ao} and a low-complexity shaping-inspired approach.
	Results show that compared to \gls{d}-\gls{ris}, \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate.
	Those benefits become more pronounced when the number of \gls{ris} elements or \gls{mimo} dimensions increase.
	Of particular interest, \gls{bd}-\gls{ris} is shown to activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at much lower transmit power than \gls{d}-\gls{ris} thanks to its singular value shaping proficiency.
\end{abstract}

\begin{IEEEkeywords}
	MIMO, RIS, channel shaping, rate maximization, singular value bounds, manifold optimization.
\end{IEEEkeywords}

\glsresetall

\begin{section}{Introduction}
	% \begin{subsection}{Background}
		Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that can serve on demand.
		This is empowered by recent advances in \gls{ris}, a programmable passive planar surface that recycles and redistributes ambient electromagnetic waves for improved wireless performance.
		A typical \gls{ris} consists of numerous low-power sub-wavelength non-resonant scattering elements, whose response can be engineered in real-time to manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019}.
		It enables low-noise full-duplex operation, featuring better flexibility than reflectarrays, lighter footprint than relays, and greater scalability than \gls{mimo} systems.
		One popular \gls{ris} research direction is \emph{joint passive and active beamforming} design with transceivers to enhance a specific performance measure, which has attracted significant interests in wireless communication \cite{Wu2019,Guo2020,Liu2022}, sensing \cite{He2022,Luo2022,Hua2023}, and power transfer literature \cite{Wu2020a,Feng2022,Zhao2022}.
		While passive beamforming at \gls{ris} suffers attenuation from double fading, it offers better asymptotic behaviors than active beamforming at transceivers (e.g., second-order array gain and fourth-order harvested power \cite{Zhao2022}).
		Another \gls{ris} application is \emph{information modulation} by periodically switching its reflection pattern within the channel coherence time.
		This creates a free-ride message stream with dual benefits: integrating with legacy transmitter for enhanced channel capacity \cite{Karasik2020,Basar2020,Ye2022}, or serving as individual source for low-power uplink communication \cite{Liang2020,Zhao2024,Yang2024}.
		Different from above, \emph{channel shaping} exploits \gls{ris} as a stand-alone device to modify the inherent properties of the wireless environment, for example, compensate for the Doppler effect \cite{Basar2021}, flatten frequency-selective channels \cite{Arslan2022}, improve \gls{mimo} channel rank \cite{Ozdogan2020a}, and artificially diversify channel over time for orthogonal \cite{Yang2019} and non-orthogonal \cite{Chen2023} multiple access schemes.
		This helps to decouple joint beamforming problems into a channel shaping stage and a conventional transceiver design stage, providing a versatile solution for various wireless applications.
	% \end{subsection}

	% \begin{subsection}{Related Works}
		At a specific resource block, channel shaping metrics can be classified into two categories.
		\begin{itemize}
			\item \emph{Singular value:} The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}. Those are closely related to performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to minor perturbations of the channel matrix;
			\item \emph{Power:} The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} in point-to-point channels and leakage interference \cite{Santamaria2023a} in interference channels. Those second-order metrics are less informative in \gls{mimo} but easier to analyze and optimize.
		\end{itemize}

		Although above works offered inspiring glimpses into the channel shaping potential of passive \gls{ris}, none attempted to disclose the entire singular value region.
		Most relevant literature \cite{ElMossallamy2021,Meng2023,Zheng2022,Huang2023,Bafghi2022,Zheng2023,Chae2023,Wu2019,Santamaria2023a} have also been limited to a \gls{d}-\gls{ris} model where each element is connected to a dedicated impedance and disconnected from others.
		As such, wave impinging on one element is entirely scattered by the same element.
		This simple architecture is modeled by a diagonal scattering matrix with unit-magnitude diagonal entries, which only applies a phase shift to the incoming signal.
		The idea was later generalized to \gls{bd}-\gls{ris} with group-connected architecture \cite{Shen2020a}, where adjacent elements within the same group are connected via passive reconfigurable components\footnote{Those components can be either symmetric (e.g., capacitors and inductors) or asymmetric (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}, resulting in symmetric and asymmetric scattering matrices, respectively.}.
		This allows wave impinging on one element to propagate within the circuit and depart partially from any element in the same group.
		It can thus manipulate both amplitude and phase of the scattered wave while remaining passive.
		Such a powerful model can be realized at reduced hardware cost using tree- and forest-connected architectures by graph theory \cite{Nerini2024}.
		\gls{bd}-\gls{ris} can also function in
		% hybrid transmitting-and-reflecting mode \cite{Li2023b} and
		multi-sector mode \cite{Li2023c} for full-space coverage and multi-user support.
		Practical challenges such as channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, and wideband modelling \cite{Li2024a} have also been studied in recent literature.
		Its beamforming effectiveness over \gls{d}-\gls{ris} and energy efficiency over active \gls{ris} and relay systems have been proved in \gls{siso} and \gls{miso} systems \cite{Shen2020a,Nerini2023,Nerini2024,Santamaria2023,Fang2023,Zhou2023,Li2023c,Soleymani2024}.
		However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} is still in the infancy stage.
		The authors of \cite{Bartoli2023} investigated the rate-optimal joint beamforming design for a fully-connected \gls{bd}-\gls{ris}-aided \gls{mimo} system where the direct channel is negligible.
		A transmitter-side \gls{bd}-\gls{ris} was introduced to
		massive \gls{mimo} systems that exploits statistical \gls{csi} for improved spectral efficiency \cite{Mishra2024}, which again assumed negligible direct channel and fully-connected \gls{bd}-\gls{ris}.
		Received power maximization with continuous-valued and discrete-valued \gls{bd}-\gls{ris} have been tackled respectively in closed form \cite{Nerini2023} and by machine learning approach \cite{Nerini2023b}, but the corresponding single-stream transceiver is rate-suboptimal.
	% \end{subsection}

	% \begin{subsection}{Contributions}
		\label{sc:contributions}
		This paper is motivated by a fundamental question:
		\emph{What is the channel shaping capability (in terms of singular values and their functions) of a passive \gls{ris} in \gls{mimo} channels?}
		Unlike existing works that focus on specific metrics or scenarios, we seek a general understanding of \gls{ris} in manipulating the wireless environment to provide insights for various wireless applications.
		% We aim for a comprehensive answer via analysis and optimization.
		The contributions are summarized below.

		First, we pioneer \gls{bd}-\gls{ris} study in general \gls{mimo} channels and interpret its shaping potential as branch matching and mode alignment.
		Branch matching refers to pairing and combining the branches (i.e., entries) of backward and forward channels associated with each \gls{bd}-\gls{ris} group.
		Mode alignment refers to aligning and ordering the modes (i.e., singular vectors) of indirect channels with those of direct channel.
		The former origins uniquely from the off-diagonal entries of the scattering matrix of \gls{bd}-\gls{ris}.

		Second, we provide an analytical answer to the shaping question under specific channel conditions.
		It is shown that \gls{bd}-\gls{ris} may achieve a larger or smaller communication \gls{dof} than \gls{d}-\gls{ris}.
		When the backward or forward channel is rank-deficient, we derive asymptotic bounds of individual singular values applying to \gls{d}- and \gls{bd}-\gls{ris}.
		When the direct channel is negligible, we recast the shaping question for fully-connected \gls{bd}-\gls{ris} as a well-studied linear algebra question and provide tight bounds (with closed-form \gls{ris} expressions) on channel singular values, power gain, and capacity.
		% The corresponding \gls{ris} scattering matrices are also derived in closed form.
		% The results are insightful for understanding the shaping advantages of \gls{bd}-\gls{ris} in \gls{mimo} channels.
		These results help to understand the fundamental limits of channel shaping and serve as a foundation for advanced \gls{ris} designs.

		% guide the design and optimization of future wireless networks

		Third, we provide a numerical answer to the shaping question by exploiting a geodesic \gls{rcg} method to optimize the \gls{bd}-\gls{ris} for a broad class of singular value functions.
		It compares favorably to legacy \gls{ris} designs in that the updates are performed multiplicatively along the shortest paths on the manifold for accelerated convergence.
		The method is then invoked in a Pareto problem to reveal the entire singular value region, which generalizes most relevant metrics and provides an intuitive shaping benchmark.

		Fourth, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem by a local-optimal \gls{ao} and a low-complexity shaping-inspired approach.
		The former iteratively updates the passive beamforming by geodesic \gls{rcg} and the active beamforming by eigenmode transmission, until convergence.
		The latter shapes the channel with \gls{ris} for maximum power gain and then performs conventional precoding.
		% Closed-form solutions are derived for the power gain maximization subproblem in different scenarios.
		% Interestingly, the rate deficit from the low-complexity design diminishes as the \gls{bd}-\gls{ris} evolves towards fully-connected.
		% We also extend both approaches to \gls{wsr} maximization in \gls{mimo} interference channels where the low-complexity design aims for minimum leakage interference instead.

		Fifth, we conduct extensive simulations to validate the analytical bounds and numerical methods.
		It is concluded that:
		\begin{itemize}
			\item \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate;
			\item A fully-connected \gls{bd}-\gls{ris} may be designed in closed-form for multi-functional optimality;
			\item The shaping advantages of \gls{bd}-\gls{ris} increase with the number of scattering elements and \gls{mimo} dimensions;
			\item \gls{bd}-\gls{ris} can activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at low transmit power thanks to its singular value shaping proficiency;
			\item The rate gap between the optimal and low-complexity beamforming designs diminishes as the \gls{ris} evolves from \gls{d} to fully-connected \gls{bd};
			\item The additional optimization cost of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} is affordable and the geodesic \gls{rcg} method is efficient on large-scale problems;
			\item The proposed \gls{ris} designs are robust to channel estimation errors and can be modified for symmetry with reasonable performance cost.
		\end{itemize}
	% \end{subsection}

	\emph{Notation:}
	Italic, bold lower-case, and bold upper-case letters indicate scalars, vectors and matrices, respectively.
	$\jmath$ denotes the imaginary unit.
	% $\mathbb{C}$ represents the set of complex numbers.
	$\mathbb{H}^{n \times n}$, $\mathbb{H}_+^{n \times n}$, and $\mathbb{U}^{n \times n}$ denote the set of $n \times n$ Hermitian, positive semi-definite, and unitary matrices, respectively.
	% $\mathbf{0}$ and $\mathbf{I}$ are the all-zero and identity matrices with appropriate size, respectively.
	$\Re\{\cdot\}$ takes the real part of a complex number.
	% $\tr(\cdot)$ and $\det(\cdot)$ evaluate the trace and determinant of a square matrix, respectively.
	$\diag(\cdot)$ constructs a square matrix with arguments on the main (block) diagonal and zeros elsewhere.
	$\sv(\cdot)$, $\ran(\cdot)$, and $\ker(\cdot)$ evaluate the singular values (in descending order), column space (range), and kernel of a matrix, respectively.
	% $\sv(\cdot)$ extracts the singular values of a matrix as a non-ascending vector.
	% $\card(\cdot)$ returns the cardinality of a set.
	$\conv(\cdot)$ returns the convex hull of arguments.
	$\mathrm{vec}(\cdot)$ stacks the columns of a matrix sequentially into a vector.
	$\lvert \cdot \rvert$, $\lVert \cdot \rVert$, and $\lVert \cdot \rVert _\mathrm{F}$ denote the absolute value, Euclidean norm, and Frobenius norm, respectively.
	$\sigma_n(\cdot)$ and $\lambda_n(\cdot)$ are the $n$-th largest singular value and eigenvalue, respectively.
	% $\boldsymbol{\sigma}(\cdot)$ and $\boldsymbol{\lambda}(\cdot)$ are the corresponding vectors.
	$(\cdot)^*$, $(\cdot)^\mathsf{T}$, $(\cdot)^\mathsf{H}$, $(\cdot)^\dagger$, $(\cdot)^{\star}$ denote the conjugate, transpose, conjugate transpose (Hermitian), Moore-Penrose inverse, and stationary point, respectively.
	$[N]$ is a shortcut for $\{1,2,\ldots,N\}$.
	$(\cdot)_{[x:y]}$ is a shortcut for $(\cdot)_x,(\cdot)_{x+1},\ldots,(\cdot)_y$.
	$\odot$ denotes the element-wise (Hadamard) product.
	$\mathcal{N}_{\mathbb{C}}(\mathbf{0}, \mathbf{\Sigma})$ is the multivariate \gls{cscg} distribution with mean $\mathbf{0}$ and covariance $\mathbf{\Sigma}$.
	$\sim$ means ``distributed as''.
\end{section}

\begin{section}{System Model}
	% \begin{subsection}{BD-RIS}
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, where group $g \in [G]$ contains $N_g$ scattering elements interconnected by real-time reconfigurable components \cite{Shen2020a} such that $N_\mathrm{S} = \sum_{g=1}^G N_g$.
		Without loss of generality we assume equal group size $N_g = L \triangleq N_\mathrm{S} / G, \ \forall g$ and no mutual coupling between elements.
		For asymmetric \gls{bd}-\gls{ris}, the overall scattering matrix is block-diagonal with unitary blocks\footnote{We assume asymmetric network by default to establish a benchmark for passive \gls{ris}. Some symmetric solutions (i.e., $\mathbf{\Theta} = \mathbf{\Theta}^\mathsf{T}$) will be discussed and their performance will be evaluated in Section~\ref{sc:simulation}.}
		\begin{equation}
			\label{eq:ris}
			\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th diagonal block modeling the response of group $g$.
		Apparently, \gls{d}-\gls{ris} is an extreme case of \eqref{eq:ris} with group size $L=1$.
		Some viable architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, \cite[Fig. 2]{Nerini2024} where the circuit topology have been modeled in $\mathbf{\Theta}$.
	% \end{subsection}

	% \begin{subsection}{MIMO Point-to-Point Channel}
		Consider a \gls{bd}-\gls{ris}-aided \gls{mimo} point-to-point channel with $N_\mathrm{T}$ and $N_\mathrm{R}$ transmit and receive antennas, respectively, and $N_\mathrm{S}$ scattering elements at the \gls{bd}-\gls{ris}.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (i.e., transmitter-receiver), backward (i.e., \gls{ris}-receiver), and forward (i.e., transmitter-\gls{ris}) channels, respectively.
		The equivalent channel depends on the \gls{bd}-\gls{ris} scattering matrix
		\begin{equation}
			\label{eq:channel}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
			%  \triangleq \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels associated with group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		% , and $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ is the indirect channel via group $g$.
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\label{eq:ris_decompose_group}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H},
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary matrices and $\mathbf{X}_g$ is a permutation matrix.
		Let $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ be the indirect channel via group $g$ and $\mathbf{H}_{\mathrm{B/F},g} = \mathbf{U}_{\mathrm{B/F},g} \mathbf{\Sigma}_{\mathrm{B/F},g} \mathbf{V}_{\mathrm{B/F},g}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel is thus
		\begin{equation}
			\label{eq:channel_svd}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\end{equation}

		\begin{remark}
			In \eqref{eq:channel_svd}, the \gls{bd}-\gls{ris} performs a blockwise unitary transformation to combine the backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels.
			These two attributes are refined respectively as:
			\begin{itemize}
				\item \emph{Branch matching:} To pair and combine the branches (i.e., entries) of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ through $\mathbf{\Theta}_g$;
				\item \emph{Mode alignment:} To align and order the modes (i.e., singular vectors) of $\{\mathbf{H}_g\}_{g \in [G]}$ with those of $\mathbf{H}_\mathrm{D}$ through $\mathbf{\Theta}$.
			\end{itemize}
		\end{remark}

		\begin{example}[\gls{siso} channel gain maximization]
			\label{eg:siso}
			Denote the direct, backward, forward channels as $h_\mathrm{D}$, $\mathbf{h}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{S} \times 1}$, and $\mathbf{h}_\mathrm{F}^\mathsf{H} \in \mathbb{C}^{1 \times N_\mathrm{S}}$, respectively.
			In this case, mode alignment boils down to phase matching and the optimal \gls{bd}-\gls{ris} structure is
			% any $L \in [N_\mathrm{S}]$, including \gls{d}-\gls{ris}, suffices for perfect mode alignment using
			\begin{equation}
				\mathbf{\Theta}_{\textnormal{P-max},g}^\textnormal{SISO} = \frac{h_\mathrm{D}}{\lvert h_\mathrm{D} \rvert} \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H}, \quad \forall g,
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr] \in \mathbb{U}^{L \times L}$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr] \in \mathbb{U}^{L \times L}$, and $\mathbf{N}_{\mathrm{B/F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of kernels of $\mathbf{h}_{\mathrm{B/F},g}$.
			It is evident that any group size $L$ (including \gls{d}-\gls{ris}) suffices for perfect phase matching.
			On the other hand, the resulting maximum channel gain depends on $L$
			\begin{equation}
				\ \lvert h \rvert = \lvert h_\mathrm{D} \rvert + \sum_g \sum_l \lvert h_{\mathrm{B},g,\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi_{\mathrm{F},g}(l)} \rvert,
			\end{equation}
			where $h_{\mathrm{B/F},g,l}$ are the $l$-th entries of $\mathbf{h}_{\mathrm{B/F},g}$, and $\pi_{\mathrm{B/F},g}$ are permutations of $[L]$ sorting their magnitude in similar orders.
			That is, the maximum \gls{siso} channel gain is attained when each \gls{bd}-\gls{ris} group, apart from phase shifting, matches the $l$-th strongest backward and forward channel branches.
			A larger $L$ provides more flexible branch matching and thus higher channel gain.
		\end{example}

		Example~\ref{eg:siso} clarifies the difference between branch matching and mode alignment and show their impacts on channel shaping.
		When it comes to \gls{mimo}, the advantage of \gls{bd}-\gls{ris} in branch matching improves since the number of available branches is proportional to $N_\mathrm{T} N_\mathrm{R}$. On the other hand, the limitation of \gls{d}-\gls{ris} in mode alignment intensifies since each element can only apply a scalar phase shift to the indirect channel of $N \triangleq \min(N_\mathrm{T}, N_\mathrm{R})$ modes.
	% \end{subsection}

	% \begin{subsection}{MIMO Interference Channel}
		% \label{sc:mimo_ic}
		% We also consider a \gls{bd}-\gls{ris}-aided \gls{mimo} interference channel of $K$ transceiver pairs where each transmitter and receiver has $N_\mathrm{T}$ and $N_\mathrm{R}$ antennas, respectively, and the \gls{bd}-\gls{ris} has $N_\mathrm{S}$ scattering elements.
		% This configuration is denoted as $(N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R})^K$.
		% Let $\mathbf{H}_\mathrm{D}^{(kj)} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B}^{(k)} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F}^{(j)} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct channel from transmitter $j$ to receiver $k$, the backward channel of receiver $k$, and the forward channel of transmitter $j$, respectively, where $(j,k) \in [K]^2$.
		% Assume all transmitter-\gls{ris}-receiver paths share the same \gls{bd}-\gls{ris} scattering matrix $\mathbf{\Theta}$.
		% The equivalent channel from transmitter $j$ to receiver $k$ is
		% \begin{equation}
		% 	\label{eq:channel_interference}
		% 	\mathbf{H}^{(kj)} = \mathbf{H}_\mathrm{D}^{(kj)} {+} \mathbf{H}_\mathrm{B}^{(k)} \mathbf{\Theta} \mathbf{H}_\mathrm{F}^{(j)} = \mathbf{H}_\mathrm{D}^{(kj)} {+} \sum_g \mathbf{H}_{\mathrm{B},g}^{(k)} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}^{(j)},
		% \end{equation}
		% where $\mathbf{H}_{\mathrm{B},g}^{(k)} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g}^{(j)} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are associated with \gls{ris} group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_{\mathrm{B}}^{(k)}$ and rows of $\mathbf{H}_{\mathrm{F}}^{(j)}$, respectively.
	% \end{subsection}
\end{section}

\begin{section}{Channel Shaping}
	In this section, we first provide an example demonstrating the \gls{mimo} channel shaping advantages of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris}, then derive some analytical bounds related to channel singular values under specific channel conditions, and provide closed-form \gls{bd}-\gls{ris} solution for several cases of interest.
	Finally, we propose a numerical method to optimize the \gls{bd}-\gls{ris} for a special class of singular value functions under general channel conditions.

>>>>

\input{preamble}
\externaldocument{main}

\makeatletter
\def\bibcite#1#2{%
	\@newl@bel{b}{#1\@extra@binfo}{%
		\hyper@@link[cite]{}{cite.#1\@extra@b@citeb}{\romannumeral 0#2}%
	}%
}
\renewcommand*{\@biblabel}[1]{[\romannumeral 0#1]}
\renewcommand{\theequation}{\roman{equation}}
\makeatother

@@
